# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r setoptions, echo=TRUE}
library(ggplot2)
#Assume the activity.csv was placed at the same directory as this Rmarkdown file.
data = read.csv("activity.csv", header=TRUE, sep=",", na.strings="NA")

#tranform date to correct Date format
data$date = as.Date(as.character(data$date), "%Y-%m-%d")
```
## What is mean total number of steps taken per day?
```{r part2}
#ignore the missing values
dataTidy = data[!is.na(data$steps), ]

## create histogram of the total number of steps taken each day
#calculate the total sum of steps w.r.t. the date
sumStep = aggregate(steps~date, dataTidy, sum)
#plot graph
ggplot(dataTidy, aes(date, steps)) + geom_bar(stat = "identity", colour = "steelblue", fill = "steelblue", width = 0.7) + labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
#disable scientific notation
options(scipen=999)
#calculate the mean and median total number of steps taken per day
meanStep = mean(sumStep$steps)
medianStep = median(sumStep$steps)
meanStep
medianStep
```


## What is the average daily activity pattern?
1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
``` {r part3}
avgStep = aggregate(steps~interval, dataTidy, mean)

#plot graph
plot(avgStep$step, type="l", xaxt="n", xlab="5-minute intervals", main="Time Series Plot of the 5-minute Interval", ylab="Average Number of Steps Taken")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxIdx = which.max(avgStep$steps)
avgStep[maxIdx, c('steps', 'interval')]
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
``` {r part4_0}
sum(is.na(data$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.  
To filling all of the missing values in the dataset, I adopted the strategy: using mean of 5-minute interval
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
``` {r part4}
#get the NA index
naIdx = which(is.na(data$steps))
naInterval = data[naIdx, 3] #get the corrosponding intervale
fillSteps = sapply(naInterval, function(x) { avgStep[(avgStep$interval==x), 2]})

#create a new dataset that is equal to the orginal dataset
# BUT with the missing data filled in
dataNew = data
dataNew[naIdx, 'steps'] = fillSteps #fill in missing data
head(dataNew)
sum(is.na(dataNew$steps))
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
ggplot(dataNew, aes(date, steps)) + geom_bar(stat = "identity", colour = "steelblue", fill = "steelblue", width = 0.7) + labs(title = "Histogram of Total Number of Steps Taken Each Day (No missing data)", x = "Date", y = "Total number of steps")
```
```{r}
#calculate the mean and median total number of steps taken per day
sumStepNew = aggregate(steps~date, dataNew, sum)
meanStepNew = mean(sumStepNew$steps)
medianStepNew = median(sumStepNew$steps)
meanStepNew
medianStepNew
```
The mean number of steps taken per day are the same (`r meanStepNew`).
But, the median number of steps taken per day are slightly different (before filling missing data: `r medianStep`, after filling missing data: `r medianStepNew`). It is probably due to filling missing values by the mean of 5 minute interval.




## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r part5}
dataNew['dateIs'] = factor(sapply(dataNew$date, function(x){ if (weekdays(x) == "Sunday" | weekdays(x) == "Saturday") { "weekend" } else { "weekday"} }))
```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
avgStepDateIs = aggregate(steps~interval + dateIs, mean, data=dataNew)

library(lattice)
xyplot( steps ~ interval | dateIs, data = avgStepDateIs, type="l", layout=c(1,2), xlab="Interval", ylab="Number of steps")

```
